---
title: "Codes_figures"
author: "Gauthier Thurin"
date: "2023-03-15"
output: html_document
---

```{r libraries}
library(plotly)
library(ggpubr)
library(ggplot2)
library(grid)
library(gridExtra)
library(latex2exp)
library(future)
plan(multisession)
mvrnorm = MASS::mvrnorm
```

```{r}
t = 8*(-1000:1000)/1000
sd = 2
sd2 = sd**2

f <- function(x){
  return( exp(-x**2/(2*sd2))/sqrt(2*pi*sd2) )
}

densite = t
i=1
for (x in t){
  densite[i] = f(x)
  i = i+1
}
  
subset = t[t>=qnorm(.95,mean=0,sd=sd)]
S=0
proba = 0
i=1
for (x in subset){
  S = S+x*f(x)
  proba = proba + f(x)
  i = i+1
}
S = S/proba
ES = -S


df = cbind(t,densite)
df = as.data.frame(df)
p <- ggplot(data=df, aes(x=t, y=densite)) + geom_point(size=0.01) + theme(axis.title = element_blank(),axis.text = element_blank(),axis.ticks = element_blank()) 
p = p + geom_segment(aes(x = qnorm(.95,mean=0,sd=sd), y = 0, xend = qnorm(.95,mean=0,sd=sd), yend = 0.12),color="blue",linetype="dashed")
p = p+ annotate("text", x = qnorm(.95,mean=0,sd=sd)+0.1, y = 0.13, label = TeX("$Q(1-\\alpha)$"),color="blue",size=4)
p = p + geom_segment(aes(x = qnorm(0.05,mean=0,sd=sd), y = 0, xend = qnorm(0.05,mean=0,sd=sd), yend = 0.12),color="blue",linetype="dashed")
p = p+ annotate("text", x = qnorm(0.05,mean=0,sd=sd), y = 0.13, label = TeX("$Q(\\alpha)$"),color="blue",size=4)
p = p + geom_segment(aes(x = S, y = 0, xend = S, yend = 0.05),color="black",linetype="dashed")
p = p+ annotate("text", x = S+0.5, y = 0.06, label = TeX("$S(1-\\alpha)$"),color="black",size=4)
p = p + geom_segment(aes(x = ES, y = 0, xend = ES, yend = 0.05),color="black",linetype="dashed")
p = p+ annotate("text", x = ES-0.3, y = 0.06, label = TeX("$E(\\alpha)$"),color="black",size=4)
p = p + geom_area(data = df[t>=qnorm(.95,mean=0,sd=sd),], fill = rgb(0, 0, 1, 0.5)) + geom_area(data = df[t<=qnorm(0.05,mean=0,sd=sd),], fill = rgb(0, 0, 1, 0.5))
p
ggsave("plot.png",height=2,width=4)
```

The computation of Monge-Kantorovich quantiles relies on the minimization of the optimal transport problem. We use a Robbins-Monro algorithm for solving OT with entropic regularization. For computational reasons, the problem of optimal transport is solved on scaled data. After having scaled `Y`, one can use the function `Robbins_Monro_algo`, to solve the EOT problem between `Y` and $U_d$. With the resulting potential `v`, one can use the functions `quantile` and `superquantile` or `expshort` to infer the objects of the same name.

One can use `transf_inv` to render the data and the quantile/superquantile points in original scale.

The next cell defines all the functions.

```{r functions for OT}
# ---
# Functions for estimating the entropic map (our regularized empirical quantile function) 
# ---

quantile<-function(eps,x,Y,v){
  # Returns the entropic map at point x with regularization parameter eps.
  exp = exp( (v-cost(x,Y))  /eps )
  pi_ = exp / sum(exp) 
  y_i_pi_i = pi_*Y
  return( colSums(y_i_pi_i) )
}

superquantile<-function(eps,x,Y,v,N=30){
  tau = norm(x,type="2")
  x = x/tau 
  t = seq(tau,1,length.out=N)
  tt = matrix(0,nrow=N,ncol=d)
  for(i in seq(1,N)){
    tt[i,] = t[i]*x
  }
  Qt <- function(u){return(quantile(eps,u,Y,v))}
  # parallelized work
  N4 = floor(N/4)
  a %<-% apply(tt[1:N4,],MARGIN=1,FUN=Qt)
  b %<-% apply(tt[(N4+1):(2*N4),],MARGIN=1,FUN=Qt)
  c %<-% apply(tt[(2*N4+1):(3*N4),],MARGIN=1,FUN=Qt)
  d %<-% apply(tt[(3*N4+1):N,],MARGIN=1,FUN=Qt)
  a = cbind(a,b,c,d)
  return(rowMeans(a))
}

expshort <-function(eps,x,Y,v,N=30){
  tau = norm(x,type="2")
  x = x/tau 
  t = seq(0,tau,length.out=N)
  tt = matrix(0,nrow=N,ncol=d)
  for(i in seq(1,N)){
    tt[i,] = t[i]*x
  }
  Qt <- function(u){return(quantile(eps,u,Y,v))}
  # parallelized work
  N4 = floor(N/4)
  a %<-% apply(tt[1:N4,],MARGIN=1,FUN=Qt)
  b %<-% apply(tt[(N4+1):(2*N4),],MARGIN=1,FUN=Qt)
  c %<-% apply(tt[(2*N4+1):(3*N4),],MARGIN=1,FUN=Qt)
  d %<-% apply(tt[(3*N4+1):N,],MARGIN=1,FUN=Qt)
  a = cbind(a,b,c,d)
  return(rowMeans(a))
}

cost<- function(x,Y){ # squared euclidean distance between x and the J points of Y. 
  normdiffx <- function(y,x0=x){ 
    toto = 0.5*norm(y-x0, type="2")**2
    return(toto)
  } #calcule la norme entre un y quelconque et le x considéré
  # apply this function to each line of Y
  res = apply(Y,MARGIN=1,FUN=normdiffx) #res must contain J elements
  return(res)
}

c_transform <- function(v, x, Y, nu, eps){ #calculate the c-transform of g
  if(eps>0){
    arg = (v-cost(x,Y))/eps
    to_sum = exp(arg)*nu
    return(-eps*log( sum(to_sum) ) )
  }else{
    cost_x = cost(x,Y)
    return(min(cost_x - v))
  }
}

grad_heps <- function(v, x, Y, nu, eps){  #calculate the gradient h_eps
  if(eps>0){
    pi = nu * exp( (v-cost(x,Y))/eps )
    return(nu-pi/sum(pi))
  }else{
    cost_x = cost(x,Y)
    j_star = which.min(cost_x - v)
    to_return = nu
    to_return[j_star] = nu[j_star] - 1
    return(to_return)
  }
}

h_eps<- function(v, x, Y, nu, eps){ #Calculate the function h_eps whose expectation equals H_eps.
  ctransf = c_transform(v,x,Y,nu,eps)
  return(sum(v*nu) + ctransf - eps)
}


Robbins_Monro_Algo <- function(Y,eps,n_iter,v0){
  #--------------
  #Function that calculate the approximation of the Wasserstein distance between (the spherical uniform) and nu (empirical measure),
  #thanks to the Robbins-Monro Stochastic Algorithm. 
  #Y denotes the support of target distribution nu, and it must have been rescaled to avoid computational issues related to entropic regularization.
  #--------------
  c = 0.51
  gamma = 1
  
  J = nrow(Y)
  d = ncol(Y)
  nu = rep(1/J,J)
  #doseur = 1/max(Y) #max(Y) must be the max of the transformed Y 
  
  v = v0
  # Robbins Monro algorithm
  for(k in seq(1,n_iter)){
    # Sample from mu
    Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
    Z = Z/norm(Z,type='2')
    r = runif(1)
    x = r*Z
    
    # Update v
    v = v + gamma/((k+1)**c) * grad_heps(v, x, Y, nu, eps) 
  }
  return(v)
}


```

# Density of our reference distribution in logarithmic scale

```{r}
# log-density of Ud
f <- function(x,y){
  if(x**2 + y**2 <= 1){
    res = log(1/sqrt(x**2+y**2))
  }else{
    res = NaN
  }
  return(res)
}

N = 100
seqx = seq(-1,1,length.out=N)
Z = matrix(-1,nrow=N,ncol=N)

for(i in 1:N){
  for(j in 1:N){
    Z[i,j] = f(seqx[i],seqx[j])
  }
}

```

```{r}
# Z is a numeric matrix that ships with R
fig <- plot_ly(x=seqx,y=seqx,z = ~Z,colors="BrBG")
fig <- fig %>% add_surface() %>% hide_colorbar()
fig
```

# Banana-shaped data

```{r}
J = 2000
d = 2

# We simulate the banana shaped distribution 
X = -1 + 2 * runif(J)
Phi = 2 * pi * runif(J)
R = 0.2 * runif(J) * (1+(1-abs(X))/2)
Y = cbind( X + R * cos(Phi) , X**2 + R * sin(Phi) )

rm(X) ; rm(R) ; rm(Phi)  #pour effacer de la mémoire

plot(Y[,1],Y[,2])

```

```{r simu banana shaped low curve}
J = 2000
d = 2

# We simulate the banana shaped distribution 
X = -1 + 2 * runif(J)
e = mvrnorm(n=J,mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
phi = 1+(3/2)*sin(X*pi/2)**2
e2 = e[,2]/1.15 + 0.5*(e[,1]**2+1.21)
X = cbind( X + 1.15*e[,1]*phi, X**2 + phi*e2 )

plot(X[,1],X[,2],col="blue")
points(Y[,1],Y[,2])
```

```{r}
# center and reduce the data Y, for computational aspects related to entropic regularization
moy = apply(Y,MARGIN=2,FUN=mean)
maxY = max(Y)
  
transf <- function(u,doseur=2){
  u = (u - moy)/ (maxY * doseur )
  return(u)
}
transf_inv <- function(u,doseur=2){
  u = u * (maxY * doseur ) + moy
  return(u)
}
Y = apply(Y,FUN=transf,MARGIN=1)
Y = t(Y)
plot(Y[,1],Y[,2],col="grey")
```

```{r}
eps = 0.005
v = runif(J)
v = v - mean(v)
v = Robbins_Monro_Algo(Y,eps=eps, n_iter=20000,v0=v)
```

After that, the center-outward quantile function can be applied to any $u\in \mathbb{B}(0,1)$.

```{r}
u = c(0.2,0.7)
Qu = quantile(eps,u,Y,v)
Qu
```

In dimension 2, one can see that this quantile function indeed characterizes the target distribution on the following plot. The next cell just compute `Q(U)` for a sample drawn from $U_d$.

```{r}
Ud = matrix(0,nrow=500,ncol=d)
for(i in seq(1,500)){
  Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
  Z = Z/norm(Z,type='2')
  Ud[i,] = runif(1)*Z
}

ImgUd = apply(Ud,MARGIN=1,function(u){return(quantile(eps,u,Y,v)) })

plot(Y[,1],Y[,2],col="grey")
points(ImgUd[1,],ImgUd[2,],col="blue")
legend("top", legend = c("Y", "Q(U)"), col = c("grey", "blue"), pch = 15, bty = "n", pt.cex = 2, cex = 0.8, horiz = TRUE, inset = c(0.1, 0.1))
```

## Contours on the unit ball

```{r}
m = 2000
Ud = matrix(0,nrow=m,ncol=d)
for(i in seq(1,m)){
  Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
  Z = Z/norm(Z,type='2')
  Ud[i,] = runif(1)*Z
}
Ud = cbind(Ud,rep("Ud",m))


ncq = 1000
sphere = matrix(0,nrow=ncq,ncol=d)
for(i in seq(1,ncq)){
  Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
  Z = Z/norm(Z,type='2')
  sphere[i,] = Z
}

N = 15
N2 = 30
angles = seq(0,2*pi,length.out=N)
data = matrix(0,nrow=N*N2,ncol=2)

for(k in 1:N){
  theta = angles[k]
  r = seq(0,1,length.out=N2)
  xs = round(r * cos(theta),3)
  ys = round(r * sin(theta),3)
  us = cbind(xs,ys )
  data[(1:N2)+(k-1)*N2,] = us
}
data = cbind(data,rep("Sign",N*N2))

rayons = c(0.1,0.3,0.5,0.7,0.9)

for(tau in rayons){
  contour = tau*sphere
  contour = cbind(contour,rep("Contours",ncq))
  data = rbind(as.matrix(data),contour)
}

data = rbind(Ud,data)
data = as.data.frame(data)
colnames(data) = c("x","y","type")
data[ , 1:2] <- apply(data[ , 1:2], MARGIN = 2,function(x) as.numeric(x))


data[ , 1:2] <- apply(data[ , 1:2], MARGIN = 2,function(x) as.numeric(x))

p = ggplot(data,mapping = aes(x=x, y=y,colour=type)) + geom_point(size=0.5) + theme(legend.position = "none",axis.text= element_blank(),axis.ticks = element_blank(),axis.title = element_blank()) + scale_color_manual(values = c("blue","red","gray") )
for(k in 1:N){
  df = data[data$type=="Sign",]
  p = p + geom_path(df[(1:N2)+(k-1)*N2,],mapping=aes(x=x, y=y,colour=type),linewidth=1) 
}
p1 = p
p
```

## Quantile contours 

```{r quantile contours}
QepsYv <- function(u){return( quantile(eps,u,Y,v) ) }
radius <- function(theta){ 
  r = seq(0,1,length.out=30)
  us = cbind(round(r * cos(theta),3), round(r * sin(theta),3) )
  return(us)
}
signcurvesQ <- function(angles){
  list_of_dfs = lapply(angles,function(theta){ return(apply(radius(theta),MARGIN=1,QepsYv) ) })
  Reduce(cbind, list_of_dfs)
}
angles = seq(0,2*pi,length.out=15) #N=15
SCQ = signcurvesQ(angles) # 2 rows, N*N2 columns
SCQ = rbind(SCQ,rep("Signs",N*N2))

ncq = 500
sphere = matrix(0,nrow=ncq,ncol=d)
angles2 = seq(0,2*pi,length.out=ncq)
for(i in seq(1,ncq)){
  theta = angles2[i]
  u = c(round(cos(theta),3),round(sin(theta),3))
  sphere[i,] = u
}
avoid_copyQ <- function(rayons){
  list_of_dfs = lapply(rayons,function(tau){ return(apply(tau*sphere,MARGIN=1,QepsYv) ) })
  return(Reduce(cbind, list_of_dfs))
} #returns an object of size (d,ncq*length(rayons)) 

len = length(rayons)
CQ <- avoid_copyQ(rayons)
CQ = rbind(CQ,rep("Contours",len*ncq))


Y2 = Y[1:2000,]

# Visualization with ggplot
SCQ = t(cbind(SCQ,CQ))
SCQ = rbind(cbind(Y2,rep("Y",nrow(Y2))),SCQ)
SCQ = as.data.frame(SCQ)
colnames(SCQ) = c("x","y","type")
SCQ[ , 1:2] <- apply(SCQ[ , 1:2], MARGIN = 2,function(x) as.numeric(x))

pQ = ggplot(SCQ,aes(x=x, y=y,colour=type)) + geom_point(size=0.5) + theme(legend.position = "none",axis.title=element_blank()) + scale_color_manual(values = c("blue","red","gray") )
for(k in 1:N){
  df = SCQ[SCQ$type=="Signs",]
  pQ = pQ + geom_path(df[(1:N2)+(k-1)*N2,],mapping=aes(x=x, y=y,colour=type),linewidth=1) 
}

pQ
#ggsave("BananeCouchee_QuantilesSignes0001.png")
```

## Expected shortfall and superquantile contours 

```{r expshort and superquantiles contours}
SepsYv <- function(u){return( superquantile(eps,u,Y,v) )}
EepsYv <- function(u){return( expshort(eps,u,Y,v) )}

##################################### 
#BEGIN TO COMPUTE SIGN CURVES IN A FORMAT THAT GGPLOT UNDERSTANDS
##################################### 
radiusE <- function(theta){  
  r = c(seq(0,0.03,length.out=5),seq(0.05,1,length.out=25))
  us = cbind(round(r * cos(theta),3), round(r * sin(theta),3) )
  return(us)
}
#different radius for expected shortfall (near from the origin)
radius <- function(theta){ 
  r = seq(0,1,length.out=30)
  us = cbind(round(r * cos(theta),3), round(r * sin(theta),3) )
  return(us)
}
signcurvesS <- function(angles){
  list_of_dfs = lapply(angles,function(theta){ return(apply(radius(theta),MARGIN=1,SepsYv) ) })
  Reduce(cbind, list_of_dfs)
}
signcurvesE <- function(angles){
  list_of_dfs = lapply(angles,function(theta){ return(apply(radiusE(theta),MARGIN=1,EepsYv) ) })
  Reduce(cbind, list_of_dfs)
}

angles = seq(0,2*pi,length.out=15) #N=15

SCS = signcurvesS(angles)
SCS = rbind(SCS,rep("Signs",N*N2)) #N2=30, nb of elements in radius

SCE = signcurvesE(angles) 
SCE = rbind(SCE,rep("Signs",N*N2)) 

#####################################
# END
##################################### 

#####################################################
## BEGIN TO COMPUTE SIGN CURVES IN A FORMAT THAT GGPLOT UNDERSTANDS 
#####################################################
ncq = 500 #size of a unit contour
sphere = matrix(0,nrow=ncq,ncol=d)
angles2 = seq(0,2*pi,length.out=ncq)
for(i in seq(1,ncq)){
  theta = angles2[i]
  u = c(round(cos(theta),3),round(sin(theta),3))
  sphere[i,] = u
}

avoid_copyS <- function(rayons){
  list_of_dfs = lapply(rayons,function(tau){ return(apply(tau*sphere,MARGIN=1,SepsYv) ) })
  return(Reduce(cbind, list_of_dfs))
}
avoid_copyE <- function(rayons){
  list_of_dfs = lapply(rayons,function(tau){ return(apply(tau*sphere,MARGIN=1,EepsYv) ) })
  return(Reduce(cbind, list_of_dfs))
}
# we use avoid_copyQ and avoid_copyS to save memory 
CS <- avoid_copyS(rayons)
CE <- avoid_copyE(rayons)

len = length(rayons)
CS = rbind(CS,rep("Contours",len*ncq))
CE = rbind(CE,rep("Contours",len*ncq))

######################################################
# END
######################################################

G = 5000
X = -1 + 2 * runif(G)
Phi = 2 * pi * runif(G)
R = 0.2 * runif(G) * (1+(1-abs(X))/2)
Y2 = cbind( X + R * cos(Phi) , X**2 + R * sin(Phi) )
moy = apply(Y2,MARGIN=2,FUN=mean)
maxY = max(Y2)
Y2 = apply(Y2,FUN=transf,MARGIN=1)
Y2 = t(Y2)

# SUPERQUANTILES
SCS = t(cbind(SCS,CS))
SCS = rbind(cbind(Y2,rep("Y",nrow(Y2))),SCS)
SCS = as.data.frame(SCS)
colnames(SCS) = c("x","y","type")
SCS[ , 1:2] <- apply(SCS[ , 1:2], MARGIN = 2,function(x) as.numeric(x))
# EXPECTED SHORTFALL
SCE = t(cbind(SCE,CE))
SCE = rbind(cbind(Y2,rep("Y",nrow(Y2))),SCE)
SCE = as.data.frame(SCE)
colnames(SCE) = c("x","y","type")
SCE[ , 1:2] <- apply(SCE[ , 1:2], MARGIN = 2,function(x) as.numeric(x))

# GRAPHICS

pS = ggplot(SCS,aes(x=x, y=y,colour=type)) + geom_point(size=1) + theme_bw()+ theme(legend.position = "none",axis.title= element_blank()) + scale_color_manual(values = c("blue","red","gray") ) 
#for(k in 1:N){
#  df = SCS[SCS$type=="Contours",]
#  pS = pS + geom_path(df[(1:N2)+(k-1)*N2,],mapping=aes(x=x, y=y,colour=type),linewidth=1) 
#}
for(k in 1:N){
  df = SCS[SCS$type=="Signs",]
  pS = pS + geom_path(df[(1:N2)+(k-1)*N2,],mapping=aes(x=x, y=y,colour=type),linewidth=1) 
}

pE = ggplot(SCE,aes(x=x, y=y,colour=type)) + geom_point(size=1) + theme_bw() + theme(legend.position = "none",axis.title= element_blank()) + scale_color_manual(values = c("blue","red","gray") )
#for(k in 1:len){
#  df = SCE[SCE$type=="Contours",]
#  pE = pE + geom_path(df[(1:ncq)+(k-1)*ncq,],mapping=aes(x=x, #y=y,colour=type),linewidth=1) 
#}
for(k in 1:N){
  df = SCE[SCE$type=="Signs",]
  pE = pE + geom_path(df[(1:N2)+(k-1)*N2,],mapping=aes(x=x, y=y,colour=type),linewidth=1) 
}
```

```{r}
pE
#ggsave("BananeCouchee_expshort0005.png",height=10,width=10)
pS
#ggsave("BananeCouchee_SuperQuantilesSignes0005.png",height=10,width=10)
```

## Graphical stuff

```{r gridExtra ie the arrow}

b = nullGrob()
mygb = function(x,y) {
  grid.bezier(x=x, y=y, gp=gpar(fill="black"), 
              arrow=arrow(type="closed", length=unit(2,"mm")))
}
pQ = pQ + theme(legend.position = "none",axis.text= element_blank(),axis.ticks = element_blank(),axis.title = element_blank())

grid.arrange(arrangeGrob(b, p1, b, heights=c(0.25,0.5,0.25)),
             b,
             arrangeGrob(b, pQ, b, heights=c(0.1,0.8, 0.1)),
             ncol=3, widths=c(0.35,0.1,0.55))

#grid.arrange(b,b,pQ,ncol=3, widths=c(0.3,0.1,0.3))

vp = viewport(x = 0.4, y=.5, width=0.09, height=0.4)
pushViewport(vp)
#grid.rect(gp=gpar(fill="black", alpha=0.1))
# Add top set of arrows
mygb(x=c(0,0.5,0.8,1), y=c(0.5,0.6,0.6,0.5))
popViewport()
#ggsave("Qmap_arrow.png",height=2,width=5)
```

# VaRs and CVaRs on simulated data

## Functions

```{r functions}

simu <- function(mean1,sigma1,mean2,sigma2,J){
  Y1 = mvrnorm(n=J,mu=mean1,Sigma=sigma1)
  Y2 = mvrnorm(n=J,mu=mean2,Sigma=sigma2)
  t = min(min(Y1),min(Y2))
  Y1 = Y1 + rep(abs(t),2)
  Y2 = Y2 + rep(abs(t),2)
  return(list(Y1,Y2))
}

max_corr <- function(Y,eps,n=500,v1,moy,maxY){
  #--------------
    # - eps is the regularization parameter of EOT and n is the number of samples drawn from Ud
  #--------------
  transf_inv <- function(u){return(u * (maxY * 2 ) + moy)}
  doseur = 1/max(Y)
  Ud = matrix(0,nrow=n,ncol=2)
  for(i in seq(1,n)){
    # Sample from mu
    Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
    Z = Z/norm(Z,type='2')
    r = runif(1)
    Ud[i,] = r*Z
  }
  # Quantiles
  QepsYv <- function(u){return( transf_inv(quantile(eps,u,Y,v1)) ) }
  TZ = apply(Ud,MARGIN=1,QepsYv) 
  res = colSums( aperm(Ud) * TZ) #contient les produits scalaires entre u et Q(u)
  return(mean(res))
}

VaR_CVaR <- function(Y,eps,tau,v1,moy,maxY){
  #--------------
  # - returns the VaRs, CVaRs (for the 2-norm )
  #- computes EOT on center-reduced, and returns the results after rescaling in original scale
  #v1 is the entropic kantorovich potential obtained through the Robbins_Monro_Algo
  #--------------
  ncq = 500
  angles = seq(0,2*pi,length.out=ncq)
  Ud = matrix(0,nrow=ncq,ncol=d)
  for(i in seq(1,ncq)){
    theta = angles[i]
    u = tau * c(round(cos(theta),3),round(sin(theta),3))
    Ud[i,] = u
  }
  transf_inv <- function(u){return(u * (maxY * 2 ) + moy)}
  # Quantiles
  QepsYv <- function(u){return( quantile(eps,u,Y,v1) ) }
  CQ1 = apply(Ud,MARGIN=1,QepsYv) 
  CQ1 = apply(CQ1, MARGIN = 2,FUN=transf_inv)
  normCQ = apply(CQ1, MARGIN = 2,sum)
  VaR = CQ1[,which.max(normCQ)] 
  # Superquantiles
  SepsYv <- function(u){return( superquantile(eps,u,Y,v1) )}
  CSQ1 = apply(Ud,MARGIN=1,SepsYv)
  CSQ1 = apply(CSQ1, MARGIN = 2,FUN=transf_inv)
  normCQ = apply(CSQ1, MARGIN = 2,sum)
  CVaR = CSQ1[,which.max(normCQ)]
  return(list(VaR,CVaR,CQ1,CSQ1))
}

calculs <- function(Y,tau,eps){
  J = nrow(Y)
  moy = apply(Y,MARGIN=2,FUN=mean)
  maxY = max(Y)
  transf <- function(u){return( (u - moy)/ (maxY * 2 ) )}
  transf_inv <- function(u){return(u * (maxY * 2 ) + moy)}
  Y = apply(Y,FUN=transf,MARGIN=1) 
  Y = t(Y)
  
  v = runif(J)
  v = v - mean(v)
  v = Robbins_Monro_Algo(Y,eps=eps, n_iter=20000,v0=v)
  
  res = VaR_CVaR(Y,eps,tau,v1=v,moy,maxY) #transf and transf_inv are used here
  VaR = res[[1]]
  CVaR = res[[2]]
  
  MC = max_corr(Y=Y,eps=eps,v1=v,moy=moy,maxY=maxY) #transf and transf_inv are used here
  risks = list(VaR,CVaR,MC) 
  
  CQ = res[[3]]
  CS = res[[4]]
  
  contours = list(CQ,CS)
  # the results are in the original scale 
  return(list(risks,contours))
}

#with the contours : 
plot_situ <- function(contours,palette,VaR, CVaR, Y,limx,limy){
  # returns a graph with Y, some contours with respective Var and CVaR
  # palette must be either 1 or 2
  # -------------------------------------
  df1 = rbind(contours[[1]],rep("Q",500))
  df2 = rbind(contours[[2]],rep("S",500))
  df1 = t(cbind(df1,df2))
  df1 = rbind(df1,cbind(Y,rep("Y",nrow(Y))))
  df1 = rbind(df1,rbind(c(VaR,"VaR"),c(CVaR,"CVaR")))
  df1 = as.data.frame(df1)
  colnames(df1) = c("x","y","type")
  df1[ , 1:2] <- apply(df1[ , 1:2], MARGIN = 2,function(x) as.numeric(x))
  
  if(palette==1){
    cbbPalette <- c("#000000", "#CC79A7", "#009E73", "#000000", "#56B4E9")
  }else{
    cbbPalette <- c("#000000", "purple", "#009E73", "#000000", "#E69F00")
  }
  p = ggplot() + theme(legend.position = "none",axis.title = element_blank()) + scale_color_manual(values = cbbPalette ) + xlim(0, limx) + ylim(0, limy)
  p = p + geom_point(df1[df1$type=="Q",],mapping=aes(x=x, y=y,colour=type),size=0.01)
  p = p + geom_point(df1[df1$type=="S",],mapping=aes(x=x, y=y,colour=type),size=0.01)
  p = p + geom_point(df1[df1$type=="Y",],mapping=aes(x=x, y=y,colour=type),size=1)
  p = p + geom_path(df1[df1$type=="Q",],mapping=aes(x=x, y=y,colour=type),linewidth=1) 
  p = p + geom_path(df1[df1$type=="S",],mapping=aes(x=x, y=y,colour=type),linewidth=1)
  p = p + geom_point(df1[df1$type=="VaR",],mapping=aes(x=x, y=y,colour=type),size=2,shape=3)
  p = p + geom_point(df1[df1$type=="CVaR",],mapping=aes(x=x, y=y,colour=type),size=2,shape=4)
  return(p)
}

#without the contours : 
plot_situ <- function(contours,palette,VaR, CVaR, Y,limx,limy){
  # returns a graph with Y, some contours with respective Var and CVaR
  # palette must be either 1 or 2
  # -------------------------------------
  df1 = cbind(Y,rep("Y",nrow(Y)))
  df1 = rbind(df1,rbind(c(VaR,"VaR"),c(CVaR,"CVaR")))
  df1 = as.data.frame(df1)
  colnames(df1) = c("x","y","type")
  df1[ , 1:2] <- apply(df1[ , 1:2], MARGIN = 2,function(x) as.numeric(x))
  
  if(palette==1){
    cbbPalette <- c("#000000", "#000000", "#56B4E9")
  }else{
    cbbPalette <- c("#000000", "#000000", "#E69F00")
  }
  p = ggplot() + theme(legend.position = "none",axis.title = element_blank()) + scale_color_manual(values = cbbPalette ) + xlim(0, limx) + ylim(0, limy)
  p = p + geom_point(df1[df1$type=="Y",],mapping=aes(x=x, y=y,colour=type),size=1,alpha=0.5)
  p = p + geom_point(df1[df1$type=="VaR",],mapping=aes(x=x, y=y,colour=type),size=2,shape=3)
  p = p + geom_point(df1[df1$type=="CVaR",],mapping=aes(x=x, y=y,colour=type),size=2,shape=4)
  return(p)
}

risks_plot <- function(Y,limx,limy,palette,tau,eps){
  #palette must be either 1 or 2 depending on Y1 or Y2
  calcs = calculs(Y,tau,eps) #here, Y is reduced for computations but not changed. Hereafter, it is in original scale
  risksY = calcs[[1]]
  contours = calcs[[2]]
  VaR = risksY[[1]] ; CVaR = risksY[[2]]
  p = plot_situ(contours,palette=palette,VaR, CVaR, Y,limx,limy)
  return(list(risksY,p))
}

final_plots <- function(Y1,Y2,tau,eps){
  # returns three plots.
  # plots 1 and 2 contain contours and VaRs and CVaRs
  # plot 3 contains the barplot
  # --------------------------
  limx = max(max(Y1[,1]),max(Y2[,1])) ; limy = max(max(Y1[,2]),max(Y2[,2]))
  L1 = risks_plot(Y=Y1,limx,limy,palette=1,tau,eps) #risks and plot for Y1 
  L2 = risks_plot(Y=Y2,limx,limy,palette=2,tau,eps) #risks and plot for Y2 
  
  risks1 = L1[[1]] # list(VaR,CVaR,MC) for situation 1
  risks2 = L2[[1]] # list(VaR,CVaR,MC) for situation 2
  
  risks1 = c(sum(risks1[[1]]),sum(risks1[[2]]),risks1[[3]]) #rhovar, rhocvar, max_corr
  risks2 = c(sum(risks2[[1]]),sum(risks2[[2]]),risks2[[3]]) #rhovar, rhocvar, max_corr
  
  #normalisation to compare the real-valued risks of Y1 and Y2 for each measure
  for(k in 1:3){
    mm = max(risks1[k],risks2[k])
    risks1[k] = risks1[k]/mm
    risks2[k] = risks2[k]/mm
  }
  
  df <- data.frame(risk=c(risks1,risks2),measure=rep(c("VaR","CVaR","MaxCor"),2),situ=c(rep("1",3),rep("2",3)))
  
  p <- ggplot(df, aes(x = measure, y = risk)) +
    geom_col(aes(color = situ, fill = situ), position = position_dodge(0.8), width = 0.7,alpha=0.5) +
    scale_color_manual(values = c("#56B4E9", "#E69F00"))+
    scale_fill_manual(values = c("#56B4E9", "#E69F00")) + theme(legend.position = "none", axis.title.x = element_blank()) + scale_x_discrete(labels=c(TeX("$\\rho^S$"),TeX("$\\rho_C$"),TeX("$\\rho^Q$")))
  return(list(L1[[2]], L2[[2]],p)) 
}

```

## Calculus for each situation


```{r}
J = 500  #number of observations for each points cloud.
d=2
eps = 0.001
tau = 0.75 #level of quantiles/superquantiles

# ---- SITUATION A
mean1 = c(1,3) ; mean2 = mean1
sigma1 = matrix(c(5,3,3,5),nrow=2,ncol=2) ; sigma2 = sigma1/5
L = simu(mean1,sigma1,mean2,sigma2,J)
Y1 = L[[1]] ; Y2 = L[[2]] #here, Y is not reduced
figureA = final_plots(Y1,Y2,tau,eps)

# ---- SITUATION B
Y1 = L[[2]]
Y2 = rbind(Y2[1:496,],matrix(c(18,17,19.8,20.2,21.1,27,22,24.3),ncol=2,nrow=4))
figureB = final_plots(Y1,Y2,tau,eps)

# ---- SITUATION C
Y1 = L[[1]]
Y2 = t(t(Y1) + c(10,0))
figureC = final_plots(Y1,Y2,tau,eps)

# ---- SITUATION D
mean1 = c(1,3) ; mean2 = mean1
sigma1 = matrix(c(3,2.5,2.5,3),nrow=2,ncol=2) 
sigma2 = matrix(c(0.2,0,0,3),nrow=2,ncol=2) 
L = simu(mean1,sigma1,mean2,sigma2,J)
Y1 = L[[1]] ; Y2 = L[[2]] #here, Y is not reduced
figureD = final_plots(Y1,Y2,tau,eps)

# ---- SITUATION E
mean1 = c(7,3) 
mean2 = c(3,7) 
sigma1 = matrix(c(3,0,0,0.5),nrow=2,ncol=2) 
sigma2 = matrix(c(0.5,0,0,3),nrow=2,ncol=2) 
L = simu(mean1,sigma1,mean2,sigma2,J)
Y1 = L[[1]] ; Y2 = L[[2]] #here, Y is not reduced
figureE = final_plots(Y1,Y2,tau,eps)

```

```{r final plots }
figureA[[1]]
figureA[[2]]
figureA[[3]]
figureB[[1]]
figureB[[2]]
figureB[[3]]
figureC[[1]]
figureC[[2]]
figureC[[3]]
figureD[[1]]
figureD[[2]]
figureD[[3]]
figureE[[1]]
figureE[[2]]
figureE[[3]]
```

```{r}
ggarrange(figureA[[1]],figureA[[2]],figureA[[3]],ncol=3,nrow=2)
ggsave("toy_ex_1.png")
ggarrange(figureB[[1]],figureB[[2]],figureB[[3]],ncol=3,nrow=2)
ggsave("toy_ex_2.png")
ggarrange(figureC[[1]],figureC[[2]],figureC[[3]],ncol=3,nrow=2)
ggsave("toy_ex_3.png")
ggarrange(figureD[[1]],figureD[[2]],figureD[[3]],ncol=3,nrow=2)
ggsave("toy_ex_4.png")
ggarrange(figureE[[1]],figureE[[2]],figureE[[3]],ncol=3,nrow=2)
ggsave("toy_ex_5.png")
```






# Q, S and E contours

```{r}
J = 2000
mean1 = c(3,-1)
sigma1 = matrix(c(2,1,1,3),nrow=2,ncol=2)
Y = mvrnorm(n=J,mu=mean1,Sigma=sigma1)

p <- ggplot(as.data.frame(Y), aes(x = V1, y =  V2))
p+geom_point()
```

```{r}
contours3 <- function(Y,eps,tau,v1,moy,maxY){
  #--------------
  # - returns the VaRs, CVaRs (for the 2-norm )
  #- computes EOT on center-reduced, and returns the results after rescaling in original scale
  #v1 is the entropic kantorovich potential obtained through the Robbins_Monro_Algo
  #--------------
  ncq = 500
  angles = seq(0,2*pi,length.out=ncq)
  Ud = matrix(0,nrow=ncq,ncol=d)
  for(i in seq(1,ncq)){
    theta = angles[i]
    u = tau * c(round(cos(theta),3),round(sin(theta),3))
    Ud[i,] = u
  }
  transf_inv <- function(u){return(u * (maxY * 2 ) + moy)}
  # Quantiles
  QepsYv <- function(u){return( quantile(eps,u,Y,v1) ) }
  CQ1 = apply(Ud,MARGIN=1,QepsYv) 
  CQ1 = apply(CQ1, MARGIN = 2,FUN=transf_inv)
  #normCQ = apply(CQ1, MARGIN = 2,sum)
  #VaR = CQ1[,which.max(normCQ)] 
  # Superquantiles
  SepsYv <- function(u){return( superquantile(eps,u,Y,v1) )}
  CSQ1 = apply(Ud,MARGIN=1,SepsYv)
  CSQ1 = apply(CSQ1, MARGIN = 2,FUN=transf_inv)
  #normCQ = apply(CSQ1, MARGIN = 2,sum)
  #CVaR = CSQ1[,which.max(normCQ)]
  # expected shortfall
  EepsYv <- function(u){return( expshort(eps,u,Y,v1) )}
  CE1 = apply(Ud,MARGIN=1,EepsYv)
  CE1 = apply(CE1, MARGIN = 2,FUN=transf_inv)
  return( list(CQ1,CSQ1,CE1) )
}
```

```{r}

tau = 0.75
eps = 0.001

# 1) OT part

moy = apply(Y,MARGIN=2,FUN=mean)
maxY = max(Y)
transf <- function(u){return( (u - moy)/ (maxY * 2 ) )}
transf_inv <- function(u){return(u * (maxY * 2 ) + moy)}
Y = apply(Y,FUN=transf,MARGIN=1)
Y = t(Y)

v = runif(J)
v = v - mean(v)
v = Robbins_Monro_Algo(Y,eps=eps, n_iter=20000,v0=v) 

# 2) compute contours
res = contours3(Y,eps,tau,v1=v,moy,maxY) 
CQ = res[[1]]
CS = res[[2]]
CE = res[[3]]

# Ud but r is 0.75 
ncq = 1000
angles = seq(0,2*pi,length.out=ncq)
Ud = matrix(0,nrow=ncq,ncol=d)
for(i in seq(1,ncq)){
  theta = angles[i]
  u = tau * c(round(cos(theta),3),round(sin(theta),3))
  Ud[i,] = u
}
CQ = apply(Ud,MARGIN=1,function(u){return(transf_inv(quantile(eps,u,Y,v))) })

# 3) ggplot stuff

df1 = rbind(CQ,rep("Q",1000))
df2 = rbind(CS,rep("S",1000))
df1 = cbind(df1,df2)
df2 = rbind(CE,rep("E",1000))
df1 = t(cbind(df1,df2))

transf_inv <- function(u){return(u * (maxY * 2 ) + moy)}

# Ud but r between 0 and 0.75
Ud = matrix(0,nrow=1000,ncol=d)
for(i in seq(1,1000)){
  Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
  Z = Z/norm(Z,type='2')
  Ud[i,] = runif(1,min=0,max=tau)*Z
}
Yplot = apply(Ud,MARGIN=1,function(u){return(transf_inv(quantile(eps,u,Y,v))) })
Yplot = rbind(Yplot,rep("Y1",nrow(Yplot)))

# Ud but r between 0.75 and 1
Ud = matrix(0,nrow=1000,ncol=d)
for(i in seq(1,1000)){
  Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
  Z = Z/norm(Z,type='2')
  Ud[i,] = runif(1,min=tau,max=1)*Z
}
toto = apply(Ud,MARGIN=1,function(u){return(transf_inv(quantile(eps,u,Y,v))) })
toto = rbind(toto,rep("Y2",nrow(toto)))
Yplot = cbind(Yplot,toto)

Yplot = t(Yplot)
df1 = rbind(df1,Yplot)
df1 = as.data.frame(df1)
colnames(df1) = c("x","y","type")
df1[ , 1:2] <- apply(df1[ , 1:2], MARGIN = 2,function(x) as.numeric(x))

cbbPalette <- c("#009E73", "blue", "red", "gray", "purple")

p = ggplot() + theme(axis.title = element_blank(),axis.text= element_blank(),axis.ticks = element_blank()) + scale_color_manual(values = cbbPalette ) 
p = p + geom_point(df1[df1$type=="Q",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="S",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="E",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="Y1",],mapping=aes(x=x, y=y,colour=type),size=1) + geom_point(df1[df1$type=="Y2",],mapping=aes(x=x, y=y,colour=type),size=1) + geom_path(df1[df1$type=="Q",],mapping=aes(x=x, y=y,colour=type),linewidth=1) + geom_path(df1[df1$type=="S",],mapping=aes(x=x, y=y,colour=type),linewidth=1) + geom_path(df1[df1$type=="E",],mapping=aes(x=x, y=y,colour=type),linewidth=1)

p
#ggsave("Troiscontours.png",height=2,width=4)
```

```{r}
cbbPalette <- c("blue", "black", "gray", rgb(0, 0, 1, 0.5))

p = ggplot() + theme(axis.title = element_blank(),axis.text= element_blank(),axis.ticks = element_blank()) + scale_color_manual(values = cbbPalette ) 
p = p + geom_point(df1[df1$type=="Q",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="S",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="Y1",],mapping=aes(x=x, y=y,colour=type),size=1) + geom_point(df1[df1$type=="Y2",],mapping=aes(x=x, y=y,colour=type),size=1) + geom_path(df1[df1$type=="Q",],mapping=aes(x=x, y=y,colour=type),linewidth=1) + geom_path(df1[df1$type=="S",],mapping=aes(x=x, y=y,colour=type),linewidth=1) 
p
```

```{r}
cbbPalette <- c("black", "blue", rgb(0, 0, 1, 0.5), "gray")

p = ggplot() + theme(legend.position = "none",axis.title = element_blank(),axis.text= element_blank(),axis.ticks = element_blank()) + scale_color_manual(values = cbbPalette ) 
p = p + geom_point(df1[df1$type=="Q",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="E",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="Y1",],mapping=aes(x=x, y=y,colour=type),size=1) + geom_point(df1[df1$type=="Y2",],mapping=aes(x=x, y=y,colour=type),size=1) + geom_path(df1[df1$type=="Q",],mapping=aes(x=x, y=y,colour=type),linewidth=1) + geom_path(df1[df1$type=="E",],mapping=aes(x=x, y=y,colour=type),linewidth=1)
p
```

```{r}

# Ud but r between 0 and 0.25
Ud = matrix(0,nrow=500,ncol=d)
sphere = matrix(0,nrow=500,ncol=d)
for(i in seq(1,500)){
  Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
  Z = Z/norm(Z,type='2')
  Ud[i,] = runif(1)*Z/4
  sphere[i,] = Z
}
Yplot = apply(Ud,MARGIN=1,function(u){return(transf_inv(quantile(eps,u,Y,v))) })
Yplot = rbind(Yplot,rep("Y1",nrow(Yplot)))

# Ud but r between 0.25 and 0.75
Ud = matrix(0,nrow=500,ncol=d)
for(i in seq(1,500)){
  Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
  Z = Z/norm(Z,type='2')
  Ud[i,] = runif(1,min=0.25,max=0.75)*Z
}
toto = apply(Ud,MARGIN=1,function(u){return(transf_inv(quantile(eps,u,Y,v))) })
toto = rbind(toto,rep("Y2",nrow(toto)))
Yplot = cbind(Yplot,toto)

# Ud but r between 0.75 and 1
Ud = matrix(0,nrow=500,ncol=d)
for(i in seq(1,500)){
  Z = mvrnorm(mu=rep(0,d),Sigma=diag(1,nrow=d,ncol=d))
  Z = Z/norm(Z,type='2')
  Ud[i,] = runif(1,min=0.75,max=1)*Z
}
toto = apply(Ud,MARGIN=1,function(u){return(transf_inv(quantile(eps,u,Y,v))) })
toto = rbind(toto,rep("Y3",nrow(toto)))
Yplot = cbind(Yplot,toto)

Yplot = t(Yplot)

# contours 
ncq = 1000
angles = seq(0,2*pi,length.out=ncq)
sphere = matrix(0,nrow=ncq,ncol=d)
for(i in seq(1,ncq)){
  theta = angles[i]
  u = c(round(cos(theta),3),round(sin(theta),3))
  sphere[i,] = u
}
CQ1 = apply(0.25*sphere,MARGIN=1,function(u){return(transf_inv(quantile(eps,u,Y,v))) })
CQ2 = apply(0.75*sphere,MARGIN=1,function(u){return(transf_inv(quantile(eps,u,Y,v))) })
df1 = rbind(CQ1,rep("CQ1",ncq))
df2 = rbind(CQ2,rep("CQ2",ncq))
df1 = cbind(df1,df2)

df1 = t(df1)

df1 = rbind(df1,Yplot)
df1 = as.data.frame(df1)
colnames(df1) = c("x","y","type")
df1[ , 1:2] <- apply(df1[ , 1:2], MARGIN = 2,function(x) as.numeric(x))
```

```{r}
cbbPalette <- c("blue", "blue", rgb(0, 0, 1, 0.5), "gray",rgb(0, 0, 1, 0.5))
p4 = ggplot() + theme(axis.title = element_blank(),axis.text= element_blank(),axis.ticks = element_blank()) + scale_color_manual(values = cbbPalette ) 
p4 = p4 + geom_point(df1[df1$type=="CQ1",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="CQ2",],mapping=aes(x=x, y=y,colour=type),size=0.01) + geom_point(df1[df1$type=="Y1",],mapping=aes(x=x, y=y,colour=type),size=1) + geom_point(df1[df1$type=="Y2",],mapping=aes(x=x, y=y,colour=type),size=1) + geom_point(df1[df1$type=="Y3",],mapping=aes(x=x, y=y,colour=type),size=1) +geom_path(df1[df1$type=="CQ1",],mapping=aes(x=x, y=y,colour=type),linewidth=1,linetype="dashed") + geom_path(df1[df1$type=="CQ2",],mapping=aes(x=x, y=y,colour=type),linewidth=1) 
p4
```
